{\bf Practical Concurrency Testing}

### 2018-10-11 thesis/defense/snakefight.tex
start writing snake fight talk

\title[Thesis Proposal]{{\bf Practical Concurrency Testing}}
\subtitle[]{a thesis proposal}
\author[Ben Blum]{Ben Blum \texttt{(bblum@cs.cmu.edu)}}

\institute[CMU CSD]{Carnegie Mellon University}
\date[]{2017, April 25}

@@

\section{Introduction}
\subsection{Motivation}

@@

	Too much manual intervention for non-expert use
	\begin{itemize}
		\item Manual kernel annotations: Landslide {\em [Blum '12]} {\tiny that's me!}
		\item State space size estimation {\em [Simsa '12]}
		\item Widespread adoption requires {\bf fully automated instrumentation}!
	\end{itemize}
	\pause
	\linegap

	Limited sources of nondeterminism
	\begin{itemize}
		\item Timer-driven schedule nondeterminism (all prior work)
		\item Weak memory nondeterminism {\em [Zhang '15]}
		\item Event-driven applications {\em [Jensen '15]}
		\item {\bf Transactional concurrency}?
	\end{itemize}
\end{frame}

@@

\begin{frame}{Thesis Statement}
	Thanks to the new algorithms, heuristics, and concurrency models I have developed,
	stateless model checking is an appropriate and accessible concurrency testing technique
	for programmers in both educational and real-world settings.
\end{frame}

@@

\section{Quicksand}

\breakslide{\Large Iterative Deepening of State Spaces}

@@

@@

\begin{frame}{Total Verification}
	\textbf{Convergence theorem:} Testing maximal state space, after detecting all data races, $\equiv$ preempting everywhere!
	\begin{itemize}
		\item Insight: $\forall$ arbitrary buggy interleaving, $\exists$ equivalent interleaving % say 'under DPOR' i guess?
			using only sync/data-race PPs.
	\end{itemize}
	\linegap

@@

	Each tested for 10 CPU-hours with
	\begin{itemize}
		\item Quicksand, finding data-race candidates with...
			\begin{itemize}
				\item ...``Pure'' Happens-Before analysis {\em [Lamport '78]}
				\item ...``Limited'' Happens-Before analysis {\em [O'Callahan '03]}
			\end{itemize}
		\item Stand-alone Landslide using ICB {\em [Musuvathi '08]}, preempting at...
			\begin{itemize}
				\item ...synchronization APIs only {\em [Simsa '10]}
				\item ...all shared memory accesses {\em [Holzmann '97]}
			\end{itemize}
	\end{itemize}
\end{frame}

@@

\begin{frame}{Deliverables}
	In the thesis, I will:
	\begin{itemize}
		\item Restate the paper {\large \smiley}
		\item Study variance of repeated Quicksand experiment trials
	\end{itemize}
\end{frame}

@@

\section{Education}

\breakslide{\Large Educational Use}

\begin{frame}{Pebbles}
	CMU (15-410)
	\linegap

	Kernel (``P3'')
	\begin{itemize}
		%\item UNIX-like educational kernel specification
		\item Students implement scheduler, synchronization, VM %from scratch %% trying not to be too arrogant ._.
		\item UNIX-like process lifecycle ({\tt fork}, {\tt exec}, {\tt exit}, {\tt wait})
		%\item Teams of 2
		\item Testing with MC requires manual annotation {\em [Blum '12]}
	\end{itemize}
	\pause
	\linegap

	Thread library (``P2'')
	\begin{itemize}
		\item POSIX-like userspace thread library %(also teams of 2)
		\item Students provided reference Pebbles implementation
		\item Fixed APIs $\rightarrow$ instrumentation can be automated
		\item May still need heuristics to handle ad-hoc synchronization
	\end{itemize}
\end{frame}

\begin{frame}{15-410 User Study}
	Method (since spring 2015)
	\begin{itemize}
		\item Guest lecture midway during P2
		\item Landslide (+ Quicksand) provided with 6 test cases and user guide
		\item Student code snapshots and bugs found recorded each use
	\end{itemize}
	\pause
	\linegap

	Results (as of spring 2016)
	\begin{itemize}
		\item 47/90 groups volunteered
		\item 85 concurrency bugs found; 53 which students then fixed
		\item 32/47 found $\ge 1$ concurrency bug, 22/47 fixed $\ge 1$
		\item No visible impact on ultimate grades...
	\end{itemize}
\end{frame}

\begin{frame}{15-410 User Study - Grades Results} % lack thereof
	\begin{center}
		\includegraphics[width=\textwidth]{../p2-p3-distribution.pdf}
	\end{center}
\end{frame}

\begin{frame}{Pintos}
	U. of Chicago (CMSC 23000), Berkeley (CS162), Stanford (CS140)
	\linegap

	Kernel (``threads'')
	\begin{itemize}
		\item Basecode provides round-robin scheduler, synchronization
		\item Students implement priority scheduling, sleep
	\end{itemize}
	\linegap

	Processes (``userprog'')
	\begin{itemize}
		\item UNIX-like process lifecycle ({\tt exec}, {\tt exit}, {\tt wait})
		\item File descriptor management ({\tt create}, {\tt remove}, {\tt open}, {\tt read}, ...)
	\end{itemize}
	\linegap

	VM, filesys projects not as concurrency-focused as above
\end{frame}

\begin{frame}{Pintos \& Landslide}
	Existing support:
	\begin{itemize}
		\item Data-race analysis
			%Lockset/vector-clock analyses
			extended to interrupt-disabling
		\item Automatically inserts (most) kernel annotations
		%\item (various engineering challenges) % palloc heap, non-direct-mapped kernelspace, skipping driver init loops
		\item Often requires manual intervention
	\end{itemize}
	% "Works enough for OOPSLA but not enough for the studence"
	\linegap

	Still needed:
	\begin{itemize}
		\item Improved automation to support corner-case student designs
		\item More concurrency test cases
		\item Remove proprietary Simics dependence by porting to Bochs
	\end{itemize}
\end{frame}

\begin{frame}{Deliverables}
	In the thesis, I will:
	\begin{itemize}
		\item Pebbles
		\begin{itemize}
			\item Discuss factors keeping Landslide from improving grades
			%\item Discuss how to make P3 and Landslide compromise to work together
			\item Study accuracy of automation heuristics
		\end{itemize}
		\item Pintos
		\begin{itemize}
			\item Open-source port of Landslide to Bochs
			\item 2 semesters of user study in Pintos class(es)
			\item Compare incidence/nature of bugs found to Pebbles
		\end{itemize}
	\end{itemize}
\end{frame}

@@

\section{Transactional Memory}
\newcommand\xbegin{\texttt{\_xbegin()}\xspace}
\newcommand\xend{\texttt{\_xend()}\xspace}
\newcommand\xabort{\texttt{\_xabort()}\xspace}

@@

\begin{frame}{Timeline}
	\begin{itemize}
		\item 2017 Apr-Aug: Port Landslide to Bochs for Pintos
		\item 2017 Apr-Aug: Transactional memory
		\item 2017 Aug-Oct: Pintos user study
		\item 2017 Oct-Nov: Pebbles user study (again)
		\item 2017 Dec:     Start writing thesis
		\item 2018 Jan-Apr: More writing \& miscellany
		\item 2018 Feb-Mar: Pintos \& Pebbles (again)
		\item 2018 Apr-May: Defend \& graduate
	\end{itemize}
\end{frame}

@@

\begin{frame}{15-410: Summary of bugs found}
	\begin{center}
	\begin{tabular}{r|cc|cc}
	%det bugs found det bugs fixed  races fixed     races found
	& \multicolumn{4}{c}{Number of groups} \\
	Number & \multicolumn{2}{c|}{Deterministic bugs} & \multicolumn{2}{c}{Concurrency bugs} \\
	of bugs & found & fixed & found & fixed \\
	\hline
	0       & 27    & 32    & 15    & 23    \\
	1       & 6     & 5     & 7     & 11    \\
	2       & 9     & 6     & 12    & 8     \\
	3       & 2     & 1     & 7     & 2     \\
	4       & 1     & 1     & 3     & 1     \\
	5       & 2     & 2     & 2     & 1     \\
	11      &       &       & 1     & 1     \\
	\hline
	%Total groups
	%       & 47    & 47    & 47    & 47    \\
	Total bugs
		& 44    & 34    & 85    & 53    \\
	\end{tabular}
	\end{center}
	%\linegap

	%E.g.: 11 groups fixed 1 bug + 8 fixed 2 + 2 $\times$ 3 + 1 $\times$ 4 + 1 $\times$ 5 + 1 overachieving group who fixed 11 = 11 + 16 + 6 + 4 + 5 + 11 = 53.
\end{frame}

\begin{frame}{HTM: Even more dimensions of nondeterminism}
	% The gotcha: When xbegin can fail, it can emit diff error codes, which depend on what happened within.
	\begin{center}
		\begin{tabular}{l}
			\texttt{if ((status = \_xbegin()) == \_XBEGIN\_STARTED) \{} \\
			\texttt{~~~~if (foo < 100)}\\
			\texttt{~~~~~~~~\_xabort();} \\
			\texttt{~~~~foo++;} \\
			\texttt{~~~~\_xend();} \\
			\texttt{\} else if (status != \_XABORT\_EXPLICIT) \{} \\
			\texttt{~~~~mutex\_lock(\&foo\_lock);} \\
			\texttt{~~~~foo++;} \\
			\texttt{~~~~mutex\_unlock(\&foo\_lock);} \\
			\texttt{\}} \\
		\end{tabular}
	\end{center}
	\linegap

	\xbegin can fail with multiple possible error codes
	\begin{itemize}
		\item MC will identify on-the-fly when (e.g.) {\tt \_XABORT\_EXPLICIT} is possible
		\item Treat such transactions as having multiple failure injection possibilities
			% XXX: How to identify when the failure status is never used? Dataflow analysis?
	\end{itemize}
\end{frame}

@@

\begin{frame}{HTM: Sources of Benchmarks}
	\begin{itemize}
		\item Shamay, https://github.com/cmpxchg16/tsx.me, 2014
		\item Dehesa-Azuara and Stanley, http://www.contrib.andrew.cmu.edu/\textasciitilde{}mdehesaa/, 2016
		\item Chapman et al., Hybrid STM/HTM for nested transactions on OpenJDK, OOPSLA 2016
		\item Wang et al., Eunomia: Scaling Concurrent Search Trees under Contention Using HTM, PPoPP 2017
	\end{itemize}
\end{frame}

@@

\begin{frame}{Don't you need \texttt{if (mutex\_held(\&foo\_lock)) \_xabort();} in that transaction on slide 31?}
	Yes.
\end{frame}

\begin{frame}{References}
	\small
	\begin{itemize}
		\item {\bf [Lamport '78]}:
			Leslie Lamport. Time, clocks, and the ordering of events in a distributed system.
			Commun. ACM, 1978.
		\item {\bf [Godefroid '97]}: Patrice Godefroid.
			VeriSoft. A tool for the automatic analysis of concurrent reactive software. CAV 1997.
		\item {\bf [Holzmann '97]}: The model checker SPIN. TOSE 1997.
		\item {\bf [Engler '03]}: Dawson Engler and Ken Ashcraft.
			RacerX: effective, static detection of race conditions and deadlocks. SOSP 2003.
		\item {\bf [O'Callahan '03]}: Robert O'Callahan and Jong-Deok Choi.
			Hybrid dynamic data race detection. PPoPP '03.
		\item {\bf [Flanagan '05]}: Cormac Flanagan and Patrice Godefroid. Dynamic partial-order reduction for
			model checking software. POPL 2005.
		\item {\bf [Musuvathi '08]}: Madanlal Musuvathi et al. Finding and Reproducing Heisenbugs in Concurrent
			Programs. OSDI 2008.
	\end{itemize}
\end{frame}
\begin{frame}{References}
	\small
	\begin{itemize}
		\item {\bf [Buckhardt '10]}: A randomized scheduler with probabilistic guarantees of finding bugs. ASPLOS 2010.
		\item {\bf [Simsa '12]}: Jiri Simsa. Runtime Estimation and Resource Allocation for
			Concurrency Testing. CMU-PDL-12-113. December 2012.
		\item {\bf [Blum '12]}: Ben Blum. Landslide: Systematic Dynamic Race Detection in Kernel
			Space. CMU-CS-12-118. May 2012.
		\item {\bf [Kasicki '12]}: Baris Kasicki. Data races vs. data race bugs: telling the difference with Portend. ASPLOS '12.
		\item {\bf [Huang '15]}: Jeff Huang. Stateless model checking concurrent programs with maximal causality reduction. PLDI '15.
		\item {\bf [Jensen '15]}: Stateless Model Checking of Event-Driven Applications. OOPSLA 2015.
	\end{itemize}
\end{frame}

>>>>

\title[]{{\bf Practical Concurrency Testing}}
\subtitle[]{or, \\ How I Learned to Stop Worrying and Love the Exponential Explosion}
\author[]{Ben Blum \texttt{(bblum@cs.cmu.edu)}}

\institute[]{Carnegie Mellon University}
\date[]{2018, October 17}

@@

\section{Part 0}
\subsection{Motivation}

@@

	Too much manual intervention for non-expert use
	\begin{itemize}
		\item Manual kernel annotations: Landslide {\em [Blum '12]} {\tiny that's me!}
			% TODO: smth about choosing pps?
		\item State space size estimation {\em [Simsa '12]}
			% TODO change this bullet pt to be more explicitly about 
		\item Widespread adoption requires {\bf fully automated instrumentation}!
	\end{itemize}
	\pause
	\linegap

	Limited sources of nondeterminism
	\begin{itemize}
		\item Timer-driven schedule nondeterminism (all prior work)
		\item Weak memory nondeterminism {\em [Zhang '15]}
		\item Event-driven applications {\em [Jensen '15]}
		\item {\bf Hardware transactional memory}?
	\end{itemize}
\end{frame}

@@

\begin{frame}{Thesis Statement}
	% data race pps (as well as txn equivalence)
	Combining both theoretically-founded automatic reduction techniques
	% iterative deepening
	% (say smth like: "landslide also has a wealth of other user-tunable heuristics
	% for expert users, which you can read about in the document; for this talk i will focus just on the CPU budget thing")
	% XXX: need to prepare for the question -- "what evidence do you have that users actually want this cpu budget stuff",
	% ",or, that it helps vs just running something unlimited vs a normal icb test"
	% (well, for the 2nd half of the Q, i can talk about how ID outperforms ICB on verficiations)
	and user-informed heuristic ones,
	stateless model checking can sufficiently mitigate exponential explosion to be
	% education
	a practical testing technique for inexperienced users
	% htm
	% TODO: this last part you just added for this talk; does the thesis version of the statmeent need to include too?
	and real-world programs alike.
\end{frame}

@@

\section{Part 1}
\subsection{Quicksand}

\breakslide{\Large Quicksand}

@@

\begin{frame}{Data Race Analysis}
	A {\bf data race} is a pair of memory accesses between two threads, where:
	\begin{itemize}
		\item At least one of the accesses is a write
		\item The threads are not holding the same mutex
		\item One access {\em happens-before} the other
			(e.g., no \texttt{cond\_signal()} in between)
	\end{itemize}
	\pause
	\linegap

	Two flavors
	\begin{itemize}
		\item {\bf Pure HB}: Whether accesses were {\em observed simultaneously}, false-negative prone
			{\em [Lamport '78]}
		\item {\bf Limited HB}: Whether threads {\em could be reordered}, false-positive prone
			{\em [O'Callahan '03]}
	\end{itemize}
\end{frame}

@@

% TODO put a pict of mario-man, i.e maximal state space

\begin{frame}{Total Verification}
	\textbf{Convergence theorem:} Testing maximal state space, after detecting all data races, $\equiv$ preempting everywhere!
	\begin{itemize}
		\item Insight: $\forall$ arbitrary buggy interleaving, $\exists$ equivalent interleaving % say 'under DPOR' i guess?
			using only sync/data-race PPs, \& Quicksand will eventually reach it.
	\end{itemize}
	\linegap

@@

	Each tested for 10 CPU-hours with
	\begin{itemize}
		\item Quicksand, finding data-race candidates with...
			\begin{itemize}
				\item ...``Pure'' Happens-Before analysis {\em [Lamport '78]}
				\item ...``Limited'' Happens-Before analysis {\em [O'Callahan '03]}
			\end{itemize}
		\item Stand-alone Landslide using ICB {\em [Musuvathi '08]}, preempting at...
			\begin{itemize}
				\item ...synchronization APIs only {\em [Simsa '10]}
				\item ...all shared memory accesses
					%{\em [Holzmann '97]} % not spin
					{\em [Yang '08]} % inspect
			\end{itemize}
	\end{itemize}
\end{frame}

@@

\begin{frame}{Summary}
	Incorporate data-race analysis to add new PPs in a feedback loop
	\linegap

	Enough time to saturate \& test every data-race PP? Full verification
	\linegap

	Otherwise? Quicksand's heuristic search ordering finds bugs faster anyway
	\pause
	\linegap

	%Future work
	%\begin{itemize}
	%	\item Extend proof to weak memory
	%	\item Incorporate ICB in Quicksand
	%	\item Improve estimation accuracy
	%\end{itemize}
\end{frame}

@@

\section{Part 2}
\subsection{Education}

\breakslide{\Large Education}

\begin{frame}{Setting}
	CMU (15-410, S'15-S'18), PSU (CMPSC 473, S'18)
	\begin{itemize}
		\item Thread library (``P2'')
			\begin{itemize}
				\item POSIX-like userspace thread library %(also teams of 2)
				\item Reference ``Pebbles'' kernel provided
					% say: later they implement this in p3
				\item Fixed APIs $\rightarrow$ instrumentation can be automated
				\item Still needs heuristics to handle ad-hoc synchronization
			\end{itemize}
			% FIXME: need introduce p3 here?
	\end{itemize}
	\pause
	\linegap

	U. Chicago (CMSC 23000, F'17)
	\begin{itemize}
		\item Pintos %(``threads'' and ``userprog'')
		\begin{itemize}
			\item Basecode provides round-robin scheduler, synchronization, VM, loader
			\item ``threads'': Students implement priority scheduling, sleep
		%\end{itemize}
		%\linegap

		%\item Processes (``userprog'')
		%\begin{itemize}
				% fork is part of exec and the loader is already provided
			\item ``userprog'': UNIX-like process lifecycle
				%({\tt exec},
				({\tt exit}, {\tt wait})
			%\item File descriptor management ({\tt create}, {\tt remove}, {\tt open}, {\tt read}, ...)
				% filesys project is beyond scope of landslide
				% DONT say: the u chicago professor chose to discard the hard concurrency parts
				% DO say, if pressured: device driver concurrency would require more engineering,
				% so we chose to limit it to these two projex
		\end{itemize}
	\end{itemize}
\end{frame}

\begin{frame}{Instrumentation}
	% Unlike kernels [Blum '12], P2 can be......
	% TODO talk abt yield loops, 
	% compare to prior technique - delay abounding - which sucks bc (verify this by reading BPOR) it doesn't actually
	% wake up the other thread at the right time (using memory conflix)

	% pintos - bunch of gross patch stuff lmao
\end{frame}

% do even need this slide?
%\begin{frame}{Evaluation Questions}
%	asfda % TODO
%	% submit better projects?
%	% get better grades?
%	% like the experience?
%\end{frame}

\begin{frame}{Method}
	P2 (CMU + PSU)
	\begin{itemize}
		\item Guest lecture midway during project
		\item Landslide (+ Quicksand) provided to students with 6 test cases %and user guide
		\item Recorded code snapshots and bugs found each use (CMU only)
		\item Surveyed students on usability, false positives, etc.
	\end{itemize}
	\pause
	\linegap

	Pintos (U. Chicago)
	\begin{itemize}
		\item TA %(Kevin Zhao)
			ran Landslide on {\em threads} and {\em userprog} submissions
		\item Returned preemption traces to students after deadline
		\item Surveyed students on ease of debugging, etc.
	\end{itemize}
	% TODO: irb stuff in a bonus slide
\end{frame}

\begin{frame}{Results}
	Participation
	\begin{itemize}
		\item CMU: 103/177 students (7 semesters)
		\item PSU: 38/136 students (1 semester)
		\item U. Chicago: 4/21 students (1 semester)
	\end{itemize}
	\linegap

	Bug-finding
	\begin{itemize}
		\item CMU: 115 deterministic, 122 concurrency bugs in 75/103 P2s
		\item U. Chicago: 3 deterministic, 4 concurrency bugs in 3/4 kernels
		% \item Case study tej
	\end{itemize}
\end{frame}

\begin{frame}{Effect on Grades}
	Compared Landslide users' vs. non-users' submitted P2 grades...
	\begin{itemize}
		\item ``Does Landslide help students submit better projects?''
	\end{itemize}
	\linegap

	...as well as subsequent P3 (kernel) grades
	\begin{itemize}
		\item ``Does Landslide teach students lasting lessons about concurrency?''
	\end{itemize}
\end{frame}

\begin{frame}{Effect on Grades}
	\begin{center}
		\includegraphics[width=\textwidth]{../photo-of-ze-studence.pdf}
	\end{center}
\end{frame}

\begin{frame}{Effect on Grades}
	\begin{center}
		\includegraphics[width=\textwidth]{../photo-of-ze-studence-cdf.pdf}
	\end{center}
\end{frame}

\begin{frame}{Statistical Significance}
	Compared users' vs. non-users' grade distributions with k-sample Anderson-Darling {\em [Anderson, Darling '52]}
	\linegap

	Control 1: non-users from {\bf same semester}
	\begin{itemize}
		\item {\bf P2} grade distributions differ with $p = 0.036 < 0.05$
		\item {\bf P3} grade distributions differ with $p = 0.034 < 0.05$
			\pause
		\item Possible bias: stronger students volunteer for Landslide in first place
	\end{itemize}
	\pause
	\linegap

	Control 2: non-users from {\bf past semesters} (when Landslide not offered)
	\begin{itemize}
		\item {\bf P2} grade distributions differ with $p = 0.671$...
		\item {\bf P3} grade distributions differ with $p = 0.145$...
			\pause
		\item Possible bias: grading criteria may drift across semesters
			% FIXME does this bullet belong in this sub list
		\item Possible bias: maybe Landslide finds bugs the official tests don't
	\end{itemize}
\end{frame}

\begin{frame}{Survey}
	14(ish) questions, investigate both education \& happiness
	\begin{itemize}
		\item {\em Education:} Did Landslide help students learn more from P2?
		\item {\em Happiness:} Did Landslide help students save time/effort?
	\end{itemize}
	\linegap

	Self-report any false positives
	\begin{itemize}
		\item Landslide crashes aside, 3 FP bug reports, all simple bugs (now fixed)
			% nonexistent MAGIC_BREAK bc of different address space in shell
			% simultaneous landslides corrupt each other
			% invalid heap access when new_pages "too close"
		\item No FPs related to yield-loop/etc heuristics
	\end{itemize}
\end{frame}

\begin{frame}{Survey Results (educational value)}
	\begin{center}
		\includegraphics[width=\textwidth]{../survey6.pdf} % easier to diagnose?
	\end{center}
\end{frame}

\begin{frame}{Survey Results (happiness value)}
	\begin{center}
		\includegraphics[width=\textwidth]{../survey5.pdf} % easy to understand?
	\end{center}
\end{frame}

\begin{frame}{Survey Results (educational value)}
	\begin{center}
		\includegraphics[width=\textwidth]{../survey9.pdf} % 100 cpu hours
	\end{center}
\end{frame}

\begin{frame}{Survey Results (educational value)}
	\begin{center}
		\includegraphics[width=\textwidth]{../survey8.pdf} % learned to understand concurrency
	\end{center}
\end{frame}

\begin{frame}{Survey Results (happiness value)}
	\begin{center}
		\includegraphics[width=\textwidth]{../survey10.pdf} % recommend to friend
	\end{center}
\end{frame}

\begin{frame}{Summary}
	Landslide helps students find \& fix bugs during projects
	\begin{itemize}
		\item 73\% of CMU users found $\ge 1$ bug
		\item 56\% of CMU users fixed $\ge 1$ bug
		\item 34\% of CMU users fixed all found bugs
	\end{itemize}
	\linegap

	Effect on overall project grades: ???
	\begin{itemize}
		\item Compensating for experimental bias difficult w/human subjects...
	\end{itemize}
	\linegap

	Survey results showed students enlightenedly happy
	\linegap

	15-410 course staff now including Landslide in official grading :)
\end{frame}

@@

\section{Part 3}
\subsection{Transactional Memory}
\newcommand\xbegin{\texttt{\_xbegin()}\xspace}
\newcommand\xend{\texttt{\_xend()}\xspace}
\newcommand\xabort{\texttt{\_xabort()}\xspace}

@@

@@

\begin{frame}{15-410...}
	% TODO:
	% study of tej/etc specific bugs
	% list tests cases provided to studence & PSU-specific ones too
\end{frame}

\begin{frame}{HTM: Even more dimensions of nondeterminism}
	% The gotcha: When xbegin can fail, it can emit diff error codes, which depend on what happened within.
	\begin{center}
		\begin{tabular}{l}
			\texttt{if ((status = \_xbegin()) == \_XBEGIN\_STARTED) \{} \\
			\texttt{~~~~if (count < 100)}\\
			\texttt{~~~~~~~~\_xabort();} \\
			\texttt{~~~~count++;} \\
			\texttt{~~~~\_xend();} \\
			\texttt{\} else if (status != \_XABORT\_EXPLICIT) \{} \\
			\texttt{~~~~mutex\_lock(\&count\_lock);} \\
			\texttt{~~~~count++;} \\
			\texttt{~~~~mutex\_unlock(\&count\_lock);} \\
			\texttt{\}} \\
		\end{tabular}
	\end{center}
	\linegap

	\xbegin can fail with multiple possible error codes
	\begin{itemize}
		\item MC will identify on-the-fly when (e.g.) {\tt \_XABORT\_EXPLICIT} is possible
		\item Treat such transactions as having multiple failure injection possibilities
		\item Future work: dataflow analysis to identify when failure reason never used
	\end{itemize}
\end{frame}

@@

\begin{frame}{HTM: Sources of Benchmarks}
	\begin{itemize}
		%\item Shamay, https://github.com/cmpxchg16/tsx.me, 2014
		\item Dehesa-Azuara and Stanley, http://www.contrib.andrew.cmu.edu/\textasciitilde{}mdehesaa/, 2016
		%\item Chapman et al., Hybrid STM/HTM for nested transactions on OpenJDK, OOPSLA 2016
		%\item Wang et al., Eunomia: Scaling Concurrent Search Trees under Contention Using HTM, PPoPP 2017
			% TODO: add, that one lock thingy
	\end{itemize}
\end{frame}

@@

% TODO: slide number c.c
\begin{frame}{Don't you need \texttt{if (mutex\_held(\&foo\_lock)) \_xabort();} in that transaction on slide 31?}
	Yes.
\end{frame}

\begin{frame}{References}
	\small
	\begin{itemize}
		\item {\bf [Lamport '78]}:
			Leslie Lamport. Time, clocks, and the ordering of events in a distributed system.
			Commun. ACM, 1978.
		\item {\bf [Godefroid '97]}: Patrice Godefroid.
			VeriSoft. A tool for the automatic analysis of concurrent reactive software. CAV 1997.
		%\item {\bf [Holzmann '97]}: The model checker SPIN. TOSE 1997.
		\item {\bf [Engler '03]}: Dawson Engler and Ken Ashcraft.
			RacerX: effective, static detection of race conditions and deadlocks. SOSP 2003.
		\item {\bf [O'Callahan '03]}: Robert O'Callahan and Jong-Deok Choi.
			Hybrid dynamic data race detection. PPoPP '03.
		\item {\bf [Flanagan '05]}: Cormac Flanagan and Patrice Godefroid. Dynamic partial-order reduction for
			model checking software. POPL 2005.
		\item {\bf [Musuvathi '08]}: Madanlal Musuvathi et al. Finding and Reproducing Heisenbugs in Concurrent
			Programs. OSDI 2008.
		\item {\bf [Yang '08]}: Yu Yang et al. Efficient stateful dynamic partial order reduction. SPIN 2008.
	\end{itemize}
\end{frame}
\begin{frame}{References}
	\small
	\begin{itemize}
		\item {\bf [Buckhardt '10]}: A randomized scheduler with probabilistic guarantees of finding bugs. ASPLOS 2010.
		\item {\bf [Simsa '12]}: Jiri Simsa. Runtime Estimation and Resource Allocation for
			Concurrency Testing. CMU-PDL-12-113. December 2012.
		\item {\bf [Blum '12]}: Ben Blum. Landslide: Systematic Dynamic Race Detection in Kernel
			Space. CMU-CS-12-118. May 2012.
		\item {\bf [Kasicki '12]}: Baris Kasicki. Data races vs. data race bugs: telling the difference with Portend. ASPLOS '12.
		\item {\bf [Huang '15]}: Jeff Huang. Stateless model checking concurrent programs with maximal causality reduction. PLDI '15.
		\item {\bf [Jensen '15]}: Stateless Model Checking of Event-Driven Applications. OOPSLA 2015.
	\end{itemize}
\end{frame}
